---
title: "p8105_hw2_zw2709"
author: zw2709
output: github_document
---

This is the Solution to p8105 Homework 2.


```{r setup}
library(tidyverse)
library(readxl)

```

## Problem 1

Read the Mr. Trashwheel dataset.

```{r}
trashwheel_df = 
  read_xlsx(
    './data/Trash-Wheel-Collection-Totals-8-6-19.xlsx',
    range = cell_cols('A:N')) %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(
    sports_balls = round(sports_balls), 
    sports_balls = as.integer(sports_balls)
  )
```


Read precipitation data for 2018 and 2017.

```{r}
precip_2018 = 
  read_excel(
    './data/Trash-Wheel-Collection-Totals-8-6-19.xlsx',
    sheet = '2018 Precipitation',
    skip = 1
  ) %>% 
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year = 2018) %>% 
  relocate(year)
```

```{r}
precip_2017 = 
  read_excel(
    './data/Trash-Wheel-Collection-Totals-8-6-19.xlsx',
    sheet = '2018 Precipitation',
    skip = 1
  ) %>% 
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year = 2017) %>% 
  relocate(year)
```

Now combine annual precipitation.

```{r}
month_df = 
  tibble(
    month = 1:12,
    month_name = month.name
  )
```

```{r}
precip_df = 
  bind_rows(precip_2018, precip_2017)
  
```

```{r}
left_join(precip_df, month_df, by = 'month')
```


This dataset contains information from the Mr. Trashwheel trash collector in Baltimore, Maryland. As trash enters the inner harbor, the trashwheel collects that trash, and stores it in a dumpster. The dataset contains information on year, month, and trash collected, including some specific kinds of trash. There are a total of `r nrow(trashwheel_df)` rows in our final dataset. Additional data sheets include month precipitation data.



